% Acknowledgments / declarations
\section*{Acknowledgments}

The authors acknowledge the Forest Resources and Ecosystem Services Hub (FRESH) at the
University of British Columbia for the research environment in which FHOPS was developed.

\section*{Funding}

Development of this work was supported by the Forest Resources and Ecosystem Services Hub
(FRESH) at the University of British Columbia, including general laboratory funding held
by Gregory Paradis. Rosalia Jaffray's graduate stipend was supported in part by the
Natural Sciences and Engineering Research Council of Canada (NSERC Discovery Grant
RGPIN-2023-04197). Kathleen Coupland's salary support was provided through the University
of British Columbia Bradshaw Research Initiative for Minerals and Mining (BRIMM), Natural
Capital and Biodiversity research theme.

\section*{Declaration of competing interest}

The authors declare that they have no known competing financial interests or personal
relationships that could have appeared to influence the work reported in this paper.

\section*{CRediT author statement}

Gregory Paradis: Conceptualization, Methodology, Software, Validation, Formal analysis,
Data curation, Writing -- original draft, Writing -- review and editing, Visualization,
Project administration. Rosalia Jaffray: Conceptualization, Methodology, Validation,
Investigation, Writing -- original draft, Writing -- review and editing. Kathleen
Coupland: Conceptualization, Methodology, Writing -- review and editing.

\section*{Data availability}

The FHOPS source code is available from the public GitHub repository and PyPI package
record listed in the software metadata tables. Manuscript build scripts, benchmark
summaries, generated tables, and figures are available from the public manuscript
repository where redistribution terms allow. Bibliographic references are cited in the
manuscript; downloaded reference documents are not redistributed publicly when copyright
or licence terms do not clearly permit it.

\section*{Declaration of generative AI and AI-assisted technologies}

During manuscript preparation, the authors used OpenAI ChatGPT, Codex, and command-line
agent tooling to assist with repository maintenance, build troubleshooting,
documentation edits, and manuscript text revision. The authors reviewed, tested, and
revised the generated output and take full responsibility for the content of this
publication.
